\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{aeguill}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tikz}
\usepackage{a4wide}



\newcommand{\strid}{\textsl{strid}}
\newcommand{\example}[1]{\begin{center}\fbox{\begin{minipage}{11cm}#1\end{minipage}}\end{center}}
\newcommand{\cinput}[1]{\vcenter{\hbox{\input{#1}}}}

\title{strid -- A string diagrams generator}
\author{Samuel Mimram}
\begin{document}
\maketitle
\tableofcontents
\newpage

\strid{} is a string diagrams generator for inclusion into \LaTeX{} files. It is
still in very $\alpha$ stage but already quite useable. It is entirely
programmed in OCaml\footnote{OCaml can be downloaded at
  \url{http://caml.inria.fr/}.}. Feel free to drop me a line at
\url{samuel.mimram@pps.jussieu.fr} if you have some comments, bug reports or
feature requests about it.

\section{Presentation of \strid{}}
\subsection{A first example}
%\begin{figure}[ht!]
\example{ Suppose that $(\mathcal{C},\otimes,I)$ is a strict monoidal
  category. A \emph{monoid} in $\mathcal{C}$ is an object $M$ of $\mathcal{C}$
  together with two maps $\mu:M\otimes M\to M$, called \emph{multiplication},
  and $\eta:I\to M$, called \emph{unit}, respectively drawn as
\[
\input{monoid_mult.tex}
\]
and
\[
\input{monoid_unit.tex}
\]
such that the equalities
\[
\cinput{monoid_assoc_l.tex}
=
\cinput{monoid_assoc_r.tex}
\]
and
\[
\cinput{monoid_neutral_l.tex}
=
\cinput{monoid_neutral.tex}
=
\cinput{monoid_neutral_r.tex}
\]
hold.
}
%\caption{Definition of monoid objects.}
%\end{figure}

Let's have a look at how we typeset the left member of the associativity
equation:
\[
\input{monoid_assoc_l.tex}
\]
The \strid{} code for this figure is
\verbatiminput{monoid_assoc_l.strid}

Despite it's apparent complexity, this code is very simple! Like every \strid{}
diagram, this code starts with ``\verb+matrix {+'' and ends with
  ``\verb+}+''. Between those lines comes the actual description of the
diagram. It is structured as a matrix whose colums are separated by ``\verb+&+''
and whose lines are separated by ``\verb+\\+''.

The rightmost multiplication is typeset by
\begin{center}
\verb+mult(ull,dl,r)+
\end{center}
Here, ``\verb+mult+'' is the kind of the operator (a multiplication-shaped one)
and its arguments specify that it should be linked to the relative positions
$(-2,1)$ (\verb+ull+ means up-left-left), $(-1,-1)$ (\verb+dl+ means down-left)
and $(1,0)$. The order in which the links should be specified is indicated on
the figure below:
\[
\input{order_mult.tex}
\]
As for other operators, links are specified inputs first and then outputs.

The labels are specified similarly by instructions like
\begin{center}
\verb+text(r)[l,t=#$M$#]+
\end{center}
This create a ``\verb+text+'' operator from here to the relative position
$(1,0)$. The brackets ``\verb+[l,t=#$M$#]+'' are here to specify optional
parameters related to this operator. The ``\verb+l+'' indicates that we are
going to add a label and the ``\verb+t=#$M$#+'' means that the label's text
should be ``\verb+$M$+''. The text between \verb+#+ is quoted uninterpreted.

Suppose that we have put the text of this figure in a file named
\verb+monoid_assoc_l.strid+. Compiling this file can be simply done by typing
\begin{center}
\verb+strid monoid_assoc_l.strid+
\end{center}
This generates a file \verb+monoid_assoc_l.tex+ which can be used in a \LaTeX{}
file like:
\begin{verbatim}
\documentclass{article}

\usepackage{tikz}

\begin{document}
\input{monoid_assoc_l.tex}
\end{document}
\end{verbatim}
You will need the \texttt{TikZ} package which can be downloaded at
\url{http://sourceforge.net/projects/pgf/}.

\bigskip

Similarly, the right member of the equation is generated in a file
\verb+monoid_assoc_r.tex+. To have the equality sign between the two diagrams
centered vertically you need to center the two diagrams. This can be done using
the \verb+\vcenter+ and \verb+\hbox+ \LaTeX{} commands as shown in the following
example:
\begin{verbatim}
\[
\vcenter{\hbox{\input{monoid_assoc_l.tex}}}
=
\vcenter{\hbox{\input{monoid_assoc_r.tex}}}
\]
\end{verbatim}

\subsection{Visualizing your diagram}
Making a nice diagram is sometimes hard and \LaTeX{} compilation of the diagrams
usually takes some time to complete. If you want to quickly see the diagram
generated by \strid{} on a file \texttt{toto.strid}, type the command
\begin{center}
  \texttt{strid -g toto.strid}
\end{center}
This will open a window in which the output diagram is displayed, which is
refreshed every time the file \texttt{toto.strid} is changed.


\section{The operators}
\subsection{Line: \texttt{line}}
\[
\input{order_line.tex}
\]
\subsection{Multiplication: \texttt{mult}}
\[
\input{order_mult.tex}
\]
\subsection{Unit: \texttt{unit}}
\[
\input{order_unit.tex}
\]
\subsection{Adjunction: \texttt{adj}}
\[
\input{order_adj.tex}
\]
\subsection{Symmetry: \texttt{sym}}
\[
\input{order_sym.tex}
\]
\subsection{Braiding: \texttt{braid}}
\[
\input{order_braid.tex}
\]
\subsection{$m,n$-ary box: \texttt{$m$box$n$}}
\[
\input{order_nboxm.tex}
\]
\subsection{Vertical box: \texttt{vbox}}
\[
\input{order_vbox.tex}
\]

\subsection{Region: \texttt{region}}
Regions can be delimited:
\[
\input{pl_composition.tex}
\]
is typeset by
\verbatiminput{pl_composition.strid}


\section{Parameters of operators}
\subsection{Labels}
Labels can be added to operators. For example the diagram
\[
\input{mult_label.tex}
\]
can be typeset by \verbatiminput{mult_label.strid} If you don't like the size of
the ellipse surrounding the label, this can of course be changed. For example,
\[
\input{mult_small_label.tex}
\]
can be typeset by
\verbatiminput{mult_small_label.strid}

Various shapes are available for labels:
\subsubsection{Triangles: \texttt{triangle} / \texttt{t}}
For example,
\[
\input{bang_maybe_cut.tex}
\]
can be typeset by
\verbatiminput{bang_maybe_cut.strid}

Here, the \texttt{s} parameter is the shape (currently, only \texttt{ellipse}
and \texttt{triangle} are available), the \texttt{d} parameter is the direction
of the triangle and the \texttt{c} parameter specifies the color of the
triangle.

\subsubsection{Rectangles: \texttt{rectangle} / \texttt{r}}
\[
\cinput{monoidal_bifunct_l.tex}
\quad=\quad
\cinput{monoidal_bifunct_r.tex}
\]
The left member is typeset by
\verbatiminput{monoidal_bifunct_l.strid}

\subsection{Arrows}
Lines can be oriented using the \texttt{a} attribute. For example,
\[
\input{mult_a.tex}
\]
can be typeset by \verbatiminput{mult_a.strid} To specify that the direction
should be backwards use the \texttt{d=b} subattribute. For example,
\[
\input{mult_ab.tex}
\]
can by typeset by \verbatiminput{mult_ab.strid}


\section{Configuration files}
All parameters can be saved in a configuration file named
\texttt{strid.conf}. To generate a configuration file, type
\begin{verbatim}
strid --dump-conf
\end{verbatim}
You can then edit \texttt{strid.conf}.

Some of the options that can be set are:
\begin{itemize}
\item \verb+line_width+: default width of a line
\item \verb+label_width+: default width of a label
\item \verb+label_height+: default height of a label
\item \verb+no_tex_environment+: do not output \verb+\begin{tikz}+ and
    \verb+\end{tikz}+
\item \verb+scaling_factor+: scale the diagrams
\item \verb+label_triangle_height+: default height of a triangular label
\item \verb+label_rectangle_width+: default width of a rectangular label
\item \verb+label_rectangle_height+: default height of a rectangular label
\item \verb+interpolation+: interpolation method for drawing lines (possible
  values are \verb+cspline+ and \verb+linear+)
\item \verb+small_circle_ray+: ray of small circles (used to tweak the drawing
  of multiplications)
\end{itemize}

\section{Examples}
\subsection{Yang-Baxter equality for braids}
\[
\cinput{yang_baxter_l.tex}
\qquad=\qquad
\cinput{yang_baxter_r.tex}
\]
Left member is typeset by
\verbatiminput{yang_baxter_l.strid}
and right member by
\verbatiminput{yang_baxter_r.strid}

\subsection{Hopf law for bialgebras}
\[
\cinput{bialgebra_hopf_l.tex}
\quad=\quad
\cinput{bialgebra_hopf_r.tex}
\]
Left member is typeset by
\verbatiminput{bialgebra_hopf_l.strid}
and right member by
\verbatiminput{bialgebra_hopf_r.strid}

\newcommand{\A}{\mathbb{A}}
\newcommand{\B}{\mathbb{B}}
\newcommand{\C}{\mathbb{C}}
\subsection{Naturality condition for natural transformations between two lax functors between bicategories}

\begin{center}
\begin{tabular}{ccc}
{\scriptsize $\input{bicat_nat_transfo_l.tex}$}
&
\qquad
\raisebox{5cm}{$=$}
\qquad
&
{\scriptsize $\input{bicat_nat_transfo_r.tex}$}
\end{tabular}
\end{center}

\noindent The code for the left-hand side of the equation is
{\scriptsize \verbatiminput{bicat_nat_transfo_l.strid}}
\noindent and the code for the right-hand side is
{\scriptsize \verbatiminput{bicat_nat_transfo_r.strid}}

\end{document}
